---
title: "Yield Trend & Stability Analyses"
author: "S. Hadasch, P. Schmidt & H.-P. Piepho"
date: "1 Februar 2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
rm(list = ls()) 
require(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
dat <- read.delim("aman.txt")
dat$Rep <- as.factor(dat$Rep)
dat$Y <- as.factor(dat$Y)
```

### Rawdata
The rawdata should be formatted like this:

```{r head(dat)}
head(dat)
```

Where $Y$ and $t_j$ both represent the year of the trial, while the former is defined as a factor and the latter as an integer variables. $x_j$ is the year the respective genotype was released, $L$ is the location, $Env$ is the year-location-combination, $G$ is the genotype, $Rep$ is the replicate, $Group$ is the ??group?? and $Yield$ is the yield.

```{r str+summary}
str(dat)
summary(dat)
```

### Stage I: genotype means per environment

```{r Import Means, include=FALSE}
dat3 <- read.delim(paste("D:/User/pschmidt/Desktop/GitHub/projectX/examplecode/Tutorial Yield Trend and Stability Analyses/aman_means.txt",sep=""))
dat3 <- dat3[,-c(2,12)]
dat3 <- dat3[order(dat3$G,dat3$Env),]
dat3$Y <- as.factor(paste(dat3$Y))
```

In order to do any yield trend and/ot yield stability analyses, genotype mean yields per environment are required.

[...]

```{r str(dat3)}
str(dat3)
```

```{r Yield - tj Plot, echo=FALSE, warning=FALSE, }
ggplot(data=dat3,
aes(x=tj, y=adjmean, colour=xj)) +
geom_jitter(width=0.25) +
scale_y_continuous(name="LS-mean: genotype per environment (adjmean)",
                   limits = c(0,max(dat3$adjmean))) +
scale_x_continuous(name="Trial Year (tj)",
                   breaks=seq(min(dat3$tj),max(dat3$tj),1)) +
theme_bw() +
theme(panel.grid.major.y = element_blank(),
      panel.grid.minor   = element_blank(),
      panel.border = element_blank(),
      axis.text.x  = element_text(angle=90,hjust=1,vjust=0.5)) +
scale_colour_gradient2("Registration\nyear (xj)",
                       midpoint = median(dat3$xj),
                       breaks = c(min(dat3$xj),
                                  median(dat3$xj),
                                  max(dat3$xj)),
                       low ="orangered3",
                       mid ="mediumspringgreen",
                       high="navy")
```

### Stage II: Shukla

```{r Shukla Model, message=FALSE, results="hide"}
require(asreml)

shuklamod <- asreml(fixed   = adjmean ~ xj + tj,
                    random  = ~ G + L + Y + L:G + Y:G + Y:L + at(G):Env, 
                    ran.order = "user",
                    weights = w,
                    family  = asreml.gaussian(dispersion=1.0),
                    control = asreml.control(pworkspace=8e10 ,
                                             workspace=8e8, 
                                             maxiter=100),
                    data    = dat3)
```

The variances can be accessed via `summary(shuklamod)$varcomp[2:3]`. After some formatting we find:

```{r Shukla VC formatting, echo=FALSE}
shukla_vc  <- summary(shuklamod)$varcomp[2:3]
  vc_names <- rownames(shukla_vc)
  vc_names <- substr(vc_names,1,regexpr('!',vc_names)-1)
  vc_names <- gsub(")"," ",vc_names)
  vc_names <- gsub("at\\(G, ","",vc_names)
shukla_vc$VC <- as.factor(vc_names)
shukla_vc$yesno <- "no"
shukla_vc[which(regexpr(":Env",shukla_vc$VC)>0),"yesno"] <- "yes"
shukla_vc$StabVar <- NA
shukla_vc[which(shukla_vc$yesno=="yes"),"StabVar"] <- gsub(" :Env","",shukla_vc[which(shukla_vc$yesno=="yes"),"VC"])                          
shukla_vc$StabVar   <- as.factor(shukla_vc$StabVar)
shukla_vc <- shukla_vc[,c(3,1,2,5)]
colnames(shukla_vc) <- c("VC","Estimate","StdErr","Variety")
shukla_out <- subset(shukla_vc, is.na(shukla_vc$Variety)==F)
shukla_out <- shukla_out[order(shukla_out$Variety),]
rownames(shukla_out) <- NULL

shukla_out <- shukla_out[,c("Variety","Estimate","StdErr")]
shukla_out[,c(2,3)] <- round(shukla_out[,c(2,3)],4)
shukla_out
```

```{r Shukla Plot, echo=FALSE}
ggplot(data=shukla_out,
       aes(x=reorder(Variety,Estimate),
           y=Estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin=Estimate-StdErr, 
                    ymax=Estimate+StdErr)) +
  theme(axis.text.x  = element_text(angle=90, hjust=1),
        axis.title.x = element_blank()) +
  scale_y_continuous(name="Shukla's stability variance with s.e.",
                     limits = c(ifelse(min(shukla_out$Estimate-shukla_out$StdErr)<0,
                                       min(shukla_out$Estimate-shukla_out$StdErr),0),
                                max(shukla_out$Estimate+shukla_out$StdErr)))
```

### Stage II: Factor-Analytic Model

```{r FA Model, message=FALSE, results="hide"}
FAmod <- asreml(fixed   = adjmean ~ xj + tj,
                random  = ~ G + L + Y + L:G + Y:G + fa(G):Env,
                ran.order = "user",
                weights = w,
                family  = asreml.gaussian(dispersion=1.0),
                control = asreml.control(pworkspace=8e10 ,
                                         workspace=8e8, 
                                         maxiter=100),
                data    = dat3)
```

The variances and lambdas can again be accessed via `summary(FAmod)$varcomp[2:3]`. After some formatting we find:

```{r FA VC formatting, echo=FALSE}
FA_vc <- summary(FAmod)$varcomp[2:3]
FA_vc$VC <- rownames(FA_vc)
rownames(FA_vc) <- NULL
FA_vc$isg   <- regexpr(":Env!G.",FA_vc$VC)
FA_vc$isfa  <- regexpr(".fa1",FA_vc$VC) 
FA_vc$group <- NA
FA_vc[which(FA_vc$isg>1) ,"group"] <- "sigma"
FA_vc[which(FA_vc$isfa>1),"group"] <- "lambda"
FA_vc[which(is.na(FA_vc$group)==T),"VC"] <- substr(FA_vc[which(is.na(FA_vc$group)==T),"VC"],1,regexpr('!',FA_vc[which(is.na(FA_vc$group)==T),"VC"])-1)
FA_vc$VC <- gsub("fa\\(G):Env!G.","",FA_vc$VC)
FA_vc$VC <- gsub(".var","",FA_vc$VC)
FA_vc$VC <- gsub(".fa1","",FA_vc$VC)
FA_vc<-FA_vc[,c(3,1,2,6)]
colnames(FA_vc) <- c("Variety","Estimate","StdErr","group")

# Transposed Output
G_var  <- subset(FA_vc, group=="sigma")[,-c(4)]
colnames(G_var) <- c("Variety","Var_Est","Var_SE")
G_lam  <- subset(FA_vc, group=="lambda")[,-c(4)]
colnames(G_lam) <- c("Variety","Lambda_Est","Lambda_SE")
FA_out <- merge(G_var, G_lam, by="Variety")
FA_out[,c(2,3,4,5)] <- round(FA_out[,c(2,3,4,5)],4)
FA_out
```
