---
title: "Yield Trend & Stability Analyses"
author: "S. Hadasch, P. Schmidt, H.-P. Piepho"
date: "1 Februar 2018"
output:
  html_document: default
---

```{r setup, include=FALSE}
rm(list = ls()) 
knitr::opts_chunk$set(echo = TRUE)
dat <- read.delim("aman.txt")
dat$Rep <- as.factor(dat$Rep)
dat$Y <- as.factor(dat$Y)

dat3 <- read.delim(paste("D:/User/pschmidt/Desktop/GitHub/projectX/examplecode/Tutorial Yield Trend and Stability Analyses/aman_means.txt",sep=""))
dat3 <- dat3[order(dat3$G,dat3$Env),]
dat3$Y <- as.factor(paste(dat3$Y))
```

## Rawdata

The rawdata should be formatted like this:

```{r}
head(dat)
```

Where $Y$ and $t_j$ both represent the year of the trial, while the former is defined as a factor and the latter as an integer variables. $x_j$ is the year the respective genotype was released, $L$ is the location, $Env$ is the year-location-combination, $G$ is the genotype, $Rep$ is the replicate, $Group$ is the ??group?? and $Yield$ is the yield.

```{r}
str(dat)
summary(dat)
```

## Stage I: genotype means per environment

In order to do any yield trend and/ot yield stability analyses, genotype mean yields per environment are required.

[...]

```{r}
str(dat3)
```

## Stage II: Shukla

```{r, message=FALSE, results="hide"}
require(asreml)

shuklamod <- asreml(fixed   = adjmean ~ xj + tj,
                    random  = ~ G + L + Y + L:G + Y:G + Y:L + at(G):Env, 
                    ran.order = "user",
                    weights = w,
                    family  = asreml.gaussian(dispersion=1.0),
                    control = asreml.control(pworkspace=8e10 ,
                                             workspace=8e8, 
                                             maxiter=100),
                    data    = dat3)
```

