---
title: 'Model Comparison Part II: Random Regressions'
author: "Hadasch, Schmidt, Piepho"
date: "05 Februar 2018"
output:
  html_document: default
---


```{r setup, include=FALSE}
require(asreml)
options(scipen=4)
#################
dataset <- "aman"
dataset <- "boro"
#################
dat3 <- read.delim(paste(dataset,"_means.txt",sep=""))
dat3 <- dat3[order(dat3$G,dat3$Env),]
dat3$Y <- as.factor(paste(dat3$Y))
```

# General Information
* Dataset = `r dataset`
* Mod1:
    + $\begin{aligned}
        var(GL)  &= \sigma_{GL}^2 \\
        var(GY)  &= \sigma_{GY}^2 \\
        var(GYL) &=           \sigma_{GYL}^2 + 
                    \sqrt{r_i}\sigma_{GYL}^2 + 
                    \sqrt{t_j}\sigma_{GYL}^2 
       \end{aligned}$  
* Mod2: 
    + $\begin{aligned}
        var(GL)  &= \sigma_{GL}^2 \\
        var(GY)  &= \sigma_{GY}^2 \\
        var(GYL) &=           \sigma_{GYL}^2 + 
                    \sqrt{r_i}\sigma_{GYL}^2 + 
                          r_i \sigma_{GYL}^2 + 
                    \sqrt{t_j}\sigma_{GYL}^2 + 
                          t_j \sigma_{GYL}^2 
       \end{aligned}$
* Mod3: 
    + $\begin{aligned}
        var(GL)  &=           \sigma_{GL}^2 + 
                    \sqrt{r_i}\sigma_{GL}^2 \\ 
        var(GY)  &=           \sigma_{GY}^2 + 
                    \sqrt{r_i}\sigma_{GY}^2 + 
                    \sqrt{t_j}\sigma_{GY}^2 \\
        var(GYL) &=           \sigma_{GYL}^2 + 
                    \sqrt{r_i}\sigma_{GYL}^2 + 
                    \sqrt{t_j}\sigma_{GYL}^2 
       \end{aligned}$
* Mod4: 
    + $\begin{aligned}
        var(GL)  &=           \sigma_{GL}^2 + 
                    \sqrt{r_i}\sigma_{GL}^2 +
                          r_i \sigma_{GL}^2 \\ 
        var(GY)  &=           \sigma_{GY}^2 + 
                    \sqrt{r_i}\sigma_{GY}^2 +
                          r_i \sigma_{GY}^2 +
                    \sqrt{t_j}\sigma_{GY}^2 +
                          t_j \sigma_{GY}^2 \\
        var(GYL) &=           \sigma_{GYL}^2 + 
                    \sqrt{r_i}\sigma_{GYL}^2 + 
                          r_i \sigma_{GYL}^2 +
                    \sqrt{t_j}\sigma_{GYL}^2 +
                          t_j \sigma_{GYL}^2
       \end{aligned}$
* Model Outputs (further formatting not shown):
    + AIC: `-2*MODELOBJECT$loglik + 2*length(MODELOBJECT$gammas)`
    + Wald test: `wald(MODELOBJECT, denDF = c("none"),ssType = c("incremental","conditional"))`
    + Variance Components: `summary(MODELOBJECT)$varcomp`
    + Residual Plot per $t_j$: `plot(dat3$tj,residuals(MODELOBJECT, type=("pearson")))`

### Create variables from ri and tj 
```{r ritj}
# "Centered" and squareroot
dat3$sc_ri <- sqrt(dat3$ri - min(dat3$ri)+1)
dat3$sc_tj <- sqrt(dat3$tj - min(dat3$tj)+1) 
# "Centered" 
dat3$c_ri  <- dat3$ri - min(dat3$ri)+1
dat3$c_tj  <- dat3$tj - min(dat3$tj)+1

# Display "centered" c_tj levels
unique(dat3$c_tj)
```


## Model 1

```{r Mod1 Gpar, include=FALSE}
Mod1 <- asreml(fixed   = adjmean ~ ri + tj,
                 random  = ~ G + L + Y + G:L + G:Y + 
                                      G:Env + 
                             sc_tj:G:Env + 
                             sc_ri:G:Env,
                 ran.order = "user",
                 weights = w,
                 family  = asreml.gaussian(dispersion=1.0),
                 control = asreml.control(pworkspace=8e10,workspace=8e8,maxiter=100),
                 data    = dat3,
                 start.values = T) 

Gpar <- Mod1$gammas.table
Gpar$Constraint[7:8] <- "U"
```

```{r Mod1, results="hide", fig.show="hide"}
Mod1 <- asreml(fixed   = adjmean ~ ri + tj,
                 random  = ~ G + L + Y + G:L + G:Y + 
                             G:Env + 
                             sc_tj:G:Env + 
                             sc_ri:G:Env,
                 ran.order = "user",
                 G.param = Gpar,
                 weights = w,
                 family  = asreml.gaussian(dispersion=1.0),
                 control = asreml.control(pworkspace=8e10,workspace=8e8,maxiter=100),
                 data    = dat3)
```

```{r Mod1 out, echo=FALSE}
-2*Mod1$loglik + 2*length(Mod1$gammas)
summary(Mod1)$varcomp[,c(2,3,5)]
wald(Mod1, denDF = c("none"),ssType = c("incremental","conditional"))[,c(1,3,4)]
plot(dat3$tj,residuals(Mod1, type=("pearson")))
```

## Model 2

```{r Mod2 Gpar, include=FALSE}
Mod2 <- asreml(fixed   = adjmean ~ ri + tj,
                 random  = ~ G + L + Y + G:L + G:Y + 
                                      G:Env + 
                             sc_tj:G:Env + c_tj:G:Env +
                             sc_ri:G:Env + c_ri:G:Env,
                 ran.order = "user",
                 weights = w,
                 family  = asreml.gaussian(dispersion=1.0),
                 control = asreml.control(pworkspace=8e10,workspace=8e8,maxiter=100),
                 data    = dat3,
                 start.values = T) 

Gpar <- Mod2$gammas.table
Gpar$Constraint[7:10] <- "U"
```

```{r Mod2, results="hide", fig.show="hide"}
Mod2 <- asreml(fixed   = adjmean ~ ri + tj,
                 random  = ~ G + L + Y + G:L + G:Y + 
                                   G:Env + 
                             sc_tj:G:Env + c_tj:G:Env +
                             sc_ri:G:Env + c_ri:G:Env,
                 ran.order = "user",
                 G.param = Gpar,
                 weights = w,
                 family  = asreml.gaussian(dispersion=1.0),
                 control = asreml.control(pworkspace=8e10,workspace=8e8,maxiter=100),
                 data    = dat3)
```

```{r Mod2 AIC}
-2* Mod2$loglik + 2*length(Mod2$gammas) # AIC
```

```{r Mod2 out, echo=FALSE}
summary(Mod2)$varcomp[,c(2,3,5)]
wald(Mod2, denDF = c("none"),ssType = c("incremental","conditional"))[,c(1,3,4)]
plot(dat3$tj,residuals(Mod2, type=("pearson")))
```

## Model 3

```{r Mod3 Gpar, include=FALSE}
Mod3 <- asreml(fixed   = adjmean ~ ri + tj,
               random  = ~ G + L + Y + 
                           G:L   + sc_ri:G:L   + 
                           G:Y   + sc_ri:G:Y   + sc_tj:G:Y   +
                           G:Env + sc_ri:G:Env + sc_tj:G:Env,
               ran.order = "user",
               weights = w,
               family  = asreml.gaussian(dispersion=1.0),
               control = asreml.control(pworkspace=8e10,workspace=8e8,maxiter=100),
               data    = dat3,
               start.values = T) 

Gpar <- Mod3$gammas.table
Gpar$Constraint[c(5,7,8,10,11)] <- "U"
```

```{r Mod3, results="hide", fig.show="hide"}
Mod3 <- asreml(fixed   = adjmean ~ ri + tj,
                 random  = ~ G + L + Y + 
                           G:L   + sc_ri:G:L   + 
                           G:Y   + sc_ri:G:Y   + sc_tj:G:Y   +
                           G:Env + sc_ri:G:Env + sc_tj:G:Env,
                 ran.order = "user",
                 G.param = Gpar,
                 weights = w,
                 family  = asreml.gaussian(dispersion=1.0),
                 control = asreml.control(pworkspace=8e10,workspace=8e8,maxiter=100),
                 data    = dat3)
```

```{r Mod3 AIC}
-2* Mod3$loglik + 2*length(Mod3$gammas) # AIC
```

```{r Mod3 out, echo=FALSE}
summary(Mod3)$varcomp[,c(2,3,5)]
wald(Mod3, denDF = c("none"),ssType = c("incremental","conditional"))[,c(1,3,4)]
plot(dat3$tj,residuals(Mod3, type=("pearson")))
```

## Model 4

```{r Mod4 Gpar, include=FALSE}
Mod4 <- asreml(fixed   = adjmean ~ ri + tj,
               random  = ~ G + L + Y + 
                           G:L   + sc_ri:G:L   + 
                                    c_ri:G:L   +
                           G:Y   + sc_ri:G:Y   + sc_tj:G:Y   +
                                    c_ri:G:Y   +  c_tj:G:Y   +
                           G:Env + sc_ri:G:Env + sc_tj:G:Env +
                                    c_ri:G:Env +  c_tj:G:Env,
               ran.order = "user",
               weights = w,
               family  = asreml.gaussian(dispersion=1.0),
               control = asreml.control(pworkspace=8e10,workspace=8e8,maxiter=100),
               data    = dat3,
               start.values = T) 

Gpar <- Mod4$gammas.table
Gpar$Constraint[c(5,6,8,9,10,11,13,14,15,16)] <- "U"
```

```{r Mod4}
Mod4 <- asreml(fixed   = adjmean ~ ri + tj,
                 random  = ~ G + L + Y + 
                           G:L   + sc_ri:G:L   + 
                                    c_ri:G:L   +
                           G:Y   + sc_ri:G:Y   + sc_tj:G:Y   +
                                    c_ri:G:Y   +  c_tj:G:Y   +
                           G:Env + sc_ri:G:Env + sc_tj:G:Env +
                                    c_ri:G:Env +  c_tj:G:Env,
                 ran.order = "user",
                 G.param = Gpar,
                 weights = w,
                 family  = asreml.gaussian(dispersion=1.0),
                 control = asreml.control(pworkspace=8e10,workspace=8e8,maxiter=100),
                 data    = dat3)
```