---
title: "Yield trend & stability analyses"
output:
  html_document: default
  pdf_document: default
---

<style>
.column-left{
  float: left;
  width: 50%;
  text-align: left;
}
.column-right{
  float: right;
  width: 50%;
  text-align: left;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
list.data    <- readRDS("list_data.rds")
list.publish <- readRDS("list_publish.rds")
library(dplyr)
library(knitr)
library(kableExtra)
options(knitr.table.format = "html")
```

## Datasets: "aman" & "boro"
Two datasets (aman, boro) were analyzed. Below, you will find a quick overview for both.

### aman
```{r dat aman, echo=FALSE}
dat <- data.table(list.data[["aman"]][[2,1]])[,-c("Env","GG")]
dat
summary(dplyr::select_if(dat, is.numeric))
Filter(Negate(is.null),lapply(dat,levels))
``` 

### boro
```{r dat boro, echo=FALSE}
dat <- data.table(list.data[["boro"]][[2,1]])[,-c("Env","GG")]
dat
summary(dplyr::select_if(dat, is.numeric))
Filter(Negate(is.null),lapply(dat,levels))
``` 

## Obtain genotype-by-environment (GxE) means
We then analyze each environment separately in order to obtain adjusted means for the respective genotypes and Smith's weights. Note that (i) we here denote an `Env` as a year-by-location combination and (ii) that we actually did the analysis not on the genotype `G`, but the combination of genotype and group (`Group`), i.e.  `GG`:

```{r analysis, eval=FALSE}
  for (sel.env in c(1:n.env)){# Loop through environments
    mod <- lm(formula = Yield ~ GG + Rep,                   # Model for RCBD
              data    = subset(dat, Env==dat$Env[sel.env])) # Data subset per Env
    m <- emmeans(mod, "GG")   # Adj. means for GG
    w <- diag(solve(vcov(m))) # Smith's weights
  }
```

We then obtain GxE means that can be plotted:

### aman
```{r GxE means aman, echo=FALSE}
list.data[["aman"]][[4,1]][,c("Env", "t.j","r.i", "GG", "adj.mean", "w")]
```

### boro
```{r GxE means boro, echo=FALSE}
list.data[["boro"]][[4,1]][,c("Env", "t.j","r.i", "GG", "adj.mean", "w")]
```

<div class="column-left">
### aman
```{r GxE means plot aman, echo=FALSE}
list.GxEplots <- readRDS("GxE plots.rds")
list.GxEplots[[1]][[1]]
list.GxEplots[[2]][[1]]
```
</div>

<div class="column-right">
### boro
```{r GxE means plot boro, echo=FALSE}
list.GxEplots[[1]][[2]]
list.GxEplots[[2]][[2]]
```
</div>

## Our models
All our models include fixed regression coefficients for the registration year $r_i$ (`ri`) and trial year $t_j$ (`tj`). Their estimates give information about the genetic and agronomic trend, respectively. In the random part of the model we have effects for genotype, year, location and their interaction effects. Note that we here write the three-way interaction `Y:L:G` as `G:Env`. Regarding their syntax, the three model denoted as "Basic", "Shukla" and "FA" deviate only in the way we define this three-way interaction effect:

+ `GG:Env` Baisc Model  
+ `at(GG):Env` Shukla Model
+ `fa(GG):Env` FA Model

Thus, e.g. the basic model would be coded as:
```{r Basic Model, eval=FALSE}
Basic <- asreml(fixed   = adj.mean ~ r.i + t.j,
                random  = ~ GG + L + Y + Y:L + L:GG + Y:GG + GG:Env, 
                weights = w,
                family  = asreml.gaussian(dispersion=1.0),
                data    = dat, 
                ran.order = "user")
```

## Estimates of regression coefficients for yield trends
Finally, we obtain estimates and tests for the genetic and agronomic trends: 

<div class="column-left">
### aman
```{r regcoeff aman, echo=FALSE}
r.aman <- list.publish[["aman"]][[1,1]]

kable(r.aman, full_width = F)  %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
``` 
</div>

<div class="column-right">
### boro
```{r regcoeff boro, echo=FALSE}
r.boro <- list.publish[["boro"]][[1,1]]

kable(r.boro, full_width = F)  %>%
  kable_styling(bootstrap_options = "striped", full_width = F, position = "left")
``` 
</div>

<div class="column-left">
```{r modfit aman, echo=FALSE}
f.aman <- list.publish[["aman"]][[3,1]]

kable(f.aman, full_width = F)  %>%
  kable_styling(bootstrap_options = "striped", full_width = F, 
                position = "left", font_size=10)
``` 
</div>

<div class="column-right">
```{r modfit boro, echo=FALSE}
f.boro <- list.publish[["boro"]][[3,1]]

kable(f.boro, full_width = F)  %>%
  kable_styling(bootstrap_options = "striped", full_width = F, 
                position = "left", font_size=10)
``` 
</div>

## Estimates of stability variances
Furthermore, we obtain stability variances for each genotype. We can try to condense the informaion gained via both, the Shukla and FA model, by plotting the results as follows:

<div class="column-left">
### aman
```{r stabvar aman, echo=FALSE}
stabplots <- readRDS("stability plots.rds")
stabplots[[1]]
kable(list.publish[[1]][[2,1]], full_width = T) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, 
                position = "left", font_size=8)
``` 
</div>

<div class="column-right">
### boro
```{r stabvar boro, echo=FALSE}
stabplots[[2]]
kable(list.publish[[2]][[2,1]], full_width = T) %>%
  kable_styling(bootstrap_options = "striped", full_width = F, 
                position = "left", font_size=8)
``` 
</div>
